Self Referencing Sequences

### 2014-11-07 draft.tex
worked on density bounds for 12

As discussed in the Section~\ref{substitutionrules}, it is unclear how to establish a set of rules for $C = (1 2)$; however, there is some structure that we can exploit. In particular, we know that the sequences $1, 1, 1$ and $2, 2, 2$ will never appear in the self-referential sequence. This is because such sequences would be part of a block of size at least $3$.

What does this mean? Well, we know that if there is a density of $1$'s and $2$'s, then that density $d_2$ is bounded between 
\[ \dfrac{1}{3} \leq d_2 \leq \dfrac{2}{3} \]

In general, this gives a procedure for computing the limiting density. For example, now we know that there must be at least one $2$ for every three numbers. So now we can look at how many $2$'s are in any given string of length $9$. We know there must be at least three $2$'s. There are only three cases where there are three 2's in a segment of $9$ numbers in the sequence, namely
\[ 2, 1, 1, 2, 1, 1 ,2 , 1, 1 \]
\[ 1, 2, 1, 1, 2, 1, 1, 2, 1 \]
\[ 1, 1, 2, 1, 1, 2, 1, 1, 2 \]
In each of these cases, we can read the subsequence that must have generated them part of them. In order, they are
\[ 2, 1, 2, 1, 2 \]
\[ 1, 2, 1, 2, 1\]
\[ 2, 1, 2, 1, 2 \]
but each of these are generated by $1, 1, 1$. This means that none of the sequences of length 9 are possible, so in addition to having one 2 for every group of 3, there is another 2 for every group of 9. Therefore, we have shown that the density must be bounded by
\[ \dfrac{1}{3} + \dfrac{1}{9} \leq d_2 \leq 1 - \dfrac{1}{3} - \dfrac{1}{9} \]
In general, we can continue this process, and we did, to show that
\[ \dfrac{1}{3} + \dfrac{1}{9} + \dfrac{1}{27} \leq d_2 \leq 1 - \dfrac{1}{3} - \dfrac{1}{9} - \dfrac{1}{27} \]
Of course, the number of cases to check increases very quickly: there are $3$ choices to put the $2$ in the groups of $3$, $3$ choices to put the 3 groups of 2's and 2 positions within these groups for the groups of 9, and so on. Its not hard to see that this number grows exponentially fast, making it computationally difficult to analyze more cases.

We believe that if there is a way to show this, we will be able to show that the density is $\frac{1}{2}$, since the density will be bounded below by $\sum_{n=1}^\infty \dfrac{1}{3^n} = \frac{1}{2}$ and bounded above by the same number. However, generalizing the procedure to improve the bound seems difficult.

>>>>

As discussed in the Section~\ref{substitutionrules}, it is unclear how to establish a set of rules for $C = (1 2)$. However, there is some structure that we can exploit. In particular, we know that the sequences $1, 1, 1$ and $2, 2, 2$ will never appear in the self-referential sequence. This is because such sequences would be part of a block of size at least $3$. There would either be a 3 earlier in the sequence, or $1,1,1$ or $2,2,2$ must be the first block. We know that neither option is possible.

What does this mean? Well, we know that if there is a limiting density of $1$'s and $2$'s, then that density $d_2$ is bounded between 
\[ \dfrac{1}{3} \leq d_2 \leq \dfrac{2}{3} \]

So now we know that there must be at least one $2$ for every three numbers. We can ask  the following similar question: how many $2$'s are in any consecutive substring of length $9$? We know there must be at least three $2$'s, but do we need more? There are only three cases where there are three 2's in a segment of $9$ numbers with every consecutive subsequence of length 3 containing at least one 2, namely
\[ 2, 1, 1, 2, 1, 1 ,2 , 1, 1 \]
\[ 1, 2, 1, 1, 2, 1, 1, 2, 1 \]
\[ 1, 1, 2, 1, 1, 2, 1, 1, 2 \]
In each of these cases, we can look at the block lengths. Since the sequence is self-referential, the consecutive subsequence of block lengths must have appeared earlier in the sequence. The block lengths for the three above sequences are:
\[ 2, 1, 2, 1, 2 \]
\[ 1, 2, 1, 2, 1\]
\[ 2, 1, 2, 1, 2 \] 
We can look at the block lengths again. If the above segments were in the self-referencing sequence, then so is $1, 1, 1$, a contradiction. This means that none of the above consecutive subsequences of length 9 are possible. This means that in every consecutive subsequence of length $9$, there must be at least $4$ 2's. In other words, in addition to having one 2 for every group of 3, there is another 2 for every group of 9. Therefore, the density must be bounded by
\[ \dfrac{1}{3} + \dfrac{1}{9} \leq d_2 \leq 1 - \dfrac{1}{3} - \dfrac{1}{9} \]
In general, we can continue this process for higher powers of $3$. We did it once more for consecutive subsequences of length 27 and found you needed thirteen 2's. Thus, the density bounded by
\[ \dfrac{1}{3} + \dfrac{1}{9} + \dfrac{1}{27} \leq d_2 \leq 1 - \dfrac{1}{3} - \dfrac{1}{9} - \dfrac{1}{27} \]
The number of cases to check increases very quickly: there are $3$ choices to put the $2$ in the groups of $3$, $3$ choices to put the 3 groups of 2's and 2 positions within these groups for the groups of 9, and so on. Its not hard to see that this number grows exponentially fast, making it computationally difficult to bound the density.

If there is a pattern which we can exploit with an inductive proof, we will be able to show that the density exists and is $\frac{1}{2}$. The density will be bounded below by $\sum_{n=1}^\infty \dfrac{1}{3^n} = \frac{1}{2}$ and bounded above by the same number. However, finding patterns in this procedure to improve the bound seems difficult.
